---
title: "Assignment_5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


Assignment 5
================


## Load packages

To start, load all the required packages with the following code.
Install them if they are not installed yet.

```{r }
library(tidyverse)
library(knitr)
library(ggplot2)
library(gapminder)
```
<br>


```{r}
housing <- read_csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/landdata_states.csv")
housing %>%
  head() %>% 
  kable() 
```

1.1 Washington DC was not assigned to a region in this dataset. According to the United States Census Bureau, however, DC is part of the South region. Here:
- Change the region of DC to “South” (Hint: there are multiple ways to do this, but mutate() and ifelse() might be helpful)
- Create a new tibble or regular dataframe consisting of this new updated region variable along with the original variables State, Date and Land.Value (and no others)
- Pull out the records from DC in this new data frame. How many records are there from DC? Show the first 6 lines.

```{r message=FALSE}
housing$region <- ifelse(housing$State == "DC" & is.na (housing$region), "South", housing$region)

housing %>%
  filter(State == "DC") %>% 
  select(region, State, Land.Value,Date) %>% 
  head(6)  %>%
  kable() 
```

##1.2

```{r message=FALSE}
housing<- housing %>% group_by(region,Date)%>%
  mutate(mean_land_value = mean(Land.Value))

housing %>% group_by(region,Date)%>%
  mutate(mean_land_value = mean(Land.Value)) %>% 
  select(region, Date, mean_land_value) %>% 
  filter(region=="Midwest")%>% 
  head(6)  %>%
  arrange(mean_land_value) %>%
  kable() 
```

##1.3

```{r, message=FALSE}
housing %>% 
  ggplot(aes(x=Date,y=mean_land_value,color=region)) +
  geom_line()
```

##2

```{r, message=FALSE}
gapminder%>%
  head()%>%
  kable()
```

##2.1
```{r message=FALSE}
gapminder %>%
  filter(year==2007)%>%
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_point()
```

##2.2
```{r message=FALSE}
gapminder %>%
  filter(year==2007)%>%
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_point()+
  geom_smooth()
```
##2.3
```{r message=FALSE}
gapminder %>%
  filter(year==2007)%>%
  filter(continent!="Oceania")%>%
  ggplot(aes(x=gdpPercap,y=lifeExp,col=continent))+
  geom_point()+
  geom_smooth( se=FALSE)
```

##2.4
```{r message=FALSE}
gapminder %>%
  filter(year==2007)%>%
  filter(continent!="Oceania")%>%
  ggplot(aes(x=gdpPercap,y=lifeExp,col=continent))+
  geom_point()+
  geom_smooth()+
  facet_wrap(.~continent)
```

##2.5
```{r message=FALSE}
gapminder %>%
  ggplot(aes(x=year, y=lifeExp, col=continent, group=country))+
  geom_line()+
  facet_wrap(.~continent)
```

##2.6
```{r message=FALSE}
gapminder %>%
  filter(continent=="Asia") %>%
  head(6)  %>%
  kable() 
```

##2.7

## Historical event: Khmer_Rouge in Cambodia in around 1980 caused 2 million people died.
```{r message=FALSE}
gapminder %>%
  filter(continent=="Asia") %>%
  ggplot(aes(x=year, y=lifeExp))+
  geom_line(col="blue")+
  facet_wrap(.~country)
```


